Global Climate of the Ultra-Hot Jupiter WASP-103\lowercase{b} from \HST\ and Spitzer Phase Curve Observations

### 2017-10-27 ms.tex
big ol' table
reran with phoenix models
spitzer systematics

We calculated the transit model $T(t)$ with the \texttt{batman} package \citep{kreidberg15a}. Many of the physical parameters are tightly constrained by \cite{southworth15}, so we fixed the orbital period, time of inferior conjunction, orbital inclination, and ratio of semi-major axis to stellar radius to the previously published values ($P = 0.925545613$ day, $t_0 = 2456836.2964455 \mathrm{BJD_{TDB}}$, $i = 87.3^\circ$, and $a/R_s = 2.999$). As a test, we fit for these parameters with the Spitzer Channel 2 light curve, which has the best phase coverage and least systematic noise of the three data sets. We found that the transit parameters are consistent with the \cite{southworth15} results, so we proceeded with the remainder of the analysis holding those parameters fixed.  The free parameters for the transit model were the transit depth $r_p$ and a linear limb darkening parameter $u$.

We calculated the planet-to-star flux $F_p/F_s$ with the \texttt{SPIDERMAN} package \citep{louden17}.  This package allows users to input a climate map (temperature or brightness as a function of latitude and longitude), and generate the corresponding flux ratio for an observation at time $t$. In our fit, we assumed the star is a FIXME PHOENIX model.  We tested three different maps: a two-temperature map, with a uniform dayside temperature $T_d$ and a uniform nightside temperature $T_n$; a map generated with spherical harmonics; and the physically-motivated kinematic model from \cite{zhang16}, which has just three free parameters (the nightside temperature $T_n$, the change in temperature from day-to-night side $\Delta_T$, and the ratio of radiative to advective timescales $\xi$).

%We assume the stellar flux $F_s(\lambda)$ is a $6110\,\mathrm{K}$ blackbody \citep{gillon14}.  (FIXME - use a PHOENIX model instead). 
We scaled the planet-to-star flux by a correction factor $c$ to account for dilution from the companion star and ellipsoidal variability due to the planet's tidal distortion. The correction factor took the form: 
\begin{equation}
	c(\lambda, t) = (1 + \alpha(\lambda))A(t)
\end{equation}
where $\alpha(\lambda)$ is the additional fractional flux from the companion star and $A(t)$ is the sky-projected area of the planet. We estimated $\alpha(\lambda)$ based on the best fit spectral energy distribution from \cite{cartier16}. The companion star contributes FIXME percent of the flux at FIXME $\mu$m.  We calculated $A(t)$ using the analytic formula from \cite{leconte11b}, equation B.9, which computes the projected area of a triaxial ellipsoid. We estimated the ellipsoid properties using Table B.3 of \cite{leconte11a}, assuming the planet radius is 1.5 $R_\mathrm{Jup}$ and age is 5 Gyr. The predicted ellipsoidal variability is shown in Figure\,\ref{fig:ellipsoidal}. At quadrature, the projected radius is $8\%$ larger than at phase zero (mid-transit). %FIXME: say this doesn't matter.
%In addition to temperature, the planet's visible surface area $[R(t)]^2$ is also expected to vary with orbital phase due to tidal distortion \citep{gillon14}. In the observed light curve, changes in planet radius are degenerate with changes in temperature.

@@

\begin{figure}
\includegraphics[width = 0.5\textwidth]{Figures/ellipsoidal.pdf}
\caption{Projected area of the planet as a function of orbital phase, normalized to unity at phase zero. The area variation was predicted analytically using the model from \citep{leconte11b}.}
\label{fig:ellipsoidal}
\end{figure}

\subsection{Systematics}
Both the \HST\ and \Spitzer\ phase curves have systematic noise caused by variations in the sensitivity of the instrument over time. For the \HST/WFC3 data, the dominant systematic is an orbit-long exponential trend due to charge traps filling up over successive exposures \citep{long15, zhu17}. For \Spitzer\, the primary source of noise is the intrapixel sensitivity effect. The detector's pixels do not have uniform sensitivity, so slight changes in telescope pointing cause the recorded flux to vary.

@@

\subsubsection{\Spitzer\ Systematics}
We fit the \Spitzer\ systematics with the BLISS mapping technique \citep{stevenson12}. This technique creates a map of the intrapixel sensitivity while simultaneously fitting for other systematics and the physical parameters of the system. In addition to the intrapixel sensitivity map, we fit the data for a linear trend in time.

\section{Results}
We achieved good quality fits!

\begin{deluxetable}{lllllll}
\tablecolumns{7}
\tablewidth{0pt}:
\tablecaption{Model Comparison \label{table:models}}
\tablehead{
\colhead{Data Set} & \colhead{Model} & \colhead{$T_\mathrm{min}$} & \colhead{$T_\mathrm{max}$} & \colhead{$\chi^2$} & \colhead{$\Delta$ AIC} & \colhead{$\Delta$ BIC}}
\startdata
WFC3 & Sph. Harmonics & 1212 & 3190 & 949.6 & 0.0 & 0.0 \\
\, & Kinematic & 1955 & 3910 & 961.6 & 12.0 & 12.0 \\
\, & Two Temp. & 0 & 2837 & 999.4 & 47.7 & 43.3 \\
Ch 1 & Sph. Harmonics & 1277 & 3442 & 0.9 & 0.0 & 0.0 \\
\, & Kinematic & 1952 & 3672 & 0.9 & 34.3 & 34.3 \\
\, & Two Temp. & 1427 & 3039 & 0.9 & 10.7 & 17.6 \\
Ch 2 & Sph. Harmonics & 930 & 3775 & 1.1 & 0.0 & 3.1 \\
\, & Kinematic & 1719 & 4257 & 1.1 & 14.35 & 17.41 \\
\, & Two Temp. & 1380 & 3305 & 1.1 & 23.9 & 0.0 \\
\enddata
\vspace{-0.8cm}
\tablecomments{comments}
\end{deluxetable}

>>>>

We calculated the transit model $T(t)$ with the \texttt{batman} package \citep{kreidberg15a}. Many of the physical parameters are tightly constrained by \cite{southworth15}, so we fixed the orbital period, time of inferior conjunction, orbital inclination, and ratio of semi-major axis to stellar radius to the previously published values ($P = 0.925545613$ day, $t_0 = 2456836.2964455\,\mathrm{BJD_{TDB}}$, $i = 87.3^\circ$, and $a/R_s = 2.999$). As a test, we fit for these parameters with the Spitzer Channel 2 light curve, which has the best phase coverage and least systematic noise of the three data sets. We found that the transit parameters are consistent with the \cite{southworth15} results, so we proceeded with the remainder of the analysis holding those parameters fixed.  The free parameters for the transit model were the transit depth $r_p$ and a linear limb darkening parameter $u$.

We calculated the planet-to-star flux $F_p/F_s$ with the \texttt{SPIDERMAN} package \citep{louden17}.  This package allows users to input a climate map (temperature or brightness as a function of latitude and longitude), and generate the corresponding flux ratio for an observation at time $t$. In our fit, we calculated the stellar flux with a PHOENIX model \citep{husser13} interpolated to an effective temperature of $6110\,\mathrm{K}$ \citep{gillon14}.  We tested three different maps: a two-temperature map, with a uniform dayside temperature $T_d$ and a uniform nightside temperature $T_n$; a map generated with spherical harmonics; and the physically-motivated kinematic model from \cite{zhang16}, which has just three free parameters (the nightside temperature $T_n$, the change in temperature from day-to-night side $\Delta_T$, and the ratio of radiative to advective timescales $\xi$).

%We assume the stellar flux $F_s(\lambda)$ is a $6110\,\mathrm{K}$ blackbody \citep{gillon14}.  (FIXME - use a PHOENIX model instead). 
We scaled the planet-to-star flux by a correction factor $c$ to account for dilution from the companion star and ellipsoidal variability due to the planet's tidal distortion. The correction factor took the form: 
\begin{equation}
	c(\lambda, t) = [1 + \alpha(\lambda)]A(t)
\end{equation}
where $\alpha(\lambda)$ is the additional fractional flux from the companion star and $A(t)$ is the sky-projected area of the planet. We estimated $\alpha(\lambda)$ based on the best fit spectral energy distribution from \cite{cartier16}. The companion star contributes FIXME percent of the flux at FIXME $\mu$m.  We calculated $A(t)$ using the analytic formula from \cite{leconte11b}, equation B.9, which computes the projected area of a triaxial ellipsoid. We estimated the ellipsoid properties using Table B.3 of \cite{leconte11a}, assuming the planet radius is 1.5 $R_\mathrm{Jup}$ and age is 5 Gyr. The predicted ellipsoidal variability is shown in Figure\,\ref{fig:ellipsoidal}. At quadrature, the projected radius is $8\%$ larger than at phase zero (mid-transit). %FIXME: say this doesn't matter.
%In addition to temperature, the planet's visible surface area $[R(t)]^2$ is also expected to vary with orbital phase due to tidal distortion \citep{gillon14}. In the observed light curve, changes in planet radius are degenerate with changes in temperature.

@@

\begin{figure}
\includegraphics[width = 0.5\textwidth]{Figures/ellipsoidal.pdf}
\caption{Projected area of the planet as a function of orbital phase, normalized to unity at phase zero. The area variation was predicted analytically using the model from \cite{leconte11b}.}
\label{fig:ellipsoidal}
\end{figure}

\subsection{Systematics}
Both the \HST\ and \Spitzer\ phase curves have systematic noise caused by variations in the sensitivity of the instrument over time. For the \HST/WFC3 data, the dominant systematic is an orbit-long exponential trend due to charge traps filling up over successive exposures \citep{long15, zhu17}. For \Spitzer\, the primary source of noise is the intrapixel sensitivity effect. The detector's pixels do not have uniform sensitivity, so slight changes in telescope pointing cause the recorded flux to vary. In Figure\,\ref{fig:systematics}, we show the raw light curves before systematic noise was removed. The systematics have comparable amplitude to the thermal phase variation signal, so they must be carefully corrected to recover the underlying planet-to-star flux.

@@

\subsubsection{\Spitzer\ Systematics}
We fit the \Spitzer\ systematics with the POET pipeline using the BLISS mapping technique \citep{stevenson12}. This approach creates a map of the intrapixel sensitivity while simultaneously fitting for other systematics and the physical parameters of the system. The sensitivity map is determined by bilinear interpolation over a grid of knots centered on the stellar flux. Each knot's sensitivity is calculated from the residuals to the light curve fit: the higher the flux values for data points near a given knot, the higher the detector sensitivity is at that position.  To avoid overfitting, we chose the grid scale such that bilinear interpolation performed better than nearest neighbor interpolation. For the $3.6\,\mu$m data, the grid scale was 0.008 pixel in both $x$ and $y$. For $4.5\,\mu$m, the scale was 0.022 pixel.

In addition to the intrapixel sensitivity variation, we fit the data for a linear trend in time. We tested a quadratic trend but did not find significant evidence for the additional model complexity based on the Bayesisan information criterion (BIC).

\section{Results}
The fitted light curves are shown in Figure\,\ref{fig:phasecurves}. This figure shows results from the kinematic model for the thermal phase variation and has instrument systematics removed.  The peak planet-to-star flux ranges from 1.5 parts-per-thousand (ppt) in the WFC3 bandpass to 5 ppt in \Spitzer Channel 2.  The planet flux changes significantly with orbital phase in all three of the data sets, suggesting a strong gradient from dayside to nightside temperature. The peak brightness occurs near phase 0.5.

\subsection{Goodness of Fit}
We performed a number of tests of the quality of the light curve fits. First, we calculated the root-mean-square (rms) of the fit residuals and compared it to the expectations for photon noise alone.

Second, we tested whether the rms decreases white noise bin size.  Figure\,\ref{fig:rms} shows the binned residuals compared to expectations for white noise (uncorrelated in time).

Third, we calculated $\chi^2$, AIC, BIC.

\begin{deluxetable*}{llllllllllll}
\tablecolumns{12}
\tablewidth{0pt}:
\tablecaption{Phase-Resolved Emission Spectra \label{table:spectra}}
\tablehead{
\colhead{$\lambda$} & \colhead{Dilution} & \colhead{0.1} & \colhead{0.2} & \colhead{0.3} & \colhead{0.4} & \colhead{0.6} & \colhead{0.7} & \colhead{0.8} & \colhead{0.9} & \colhead{X} & \colhead{Y}}
\startdata
1.175 & 0.10 & $ 179 \pm 79 $ & $ 411 \pm 77 $ & $ 647 \pm 80 $ & $ 1143 \pm 65 $ & $ -15 \pm 66 $ & $ 142 \pm 67 $ & $ 1063 \pm 64 $ & $ 710 \pm 73 $ & $ 412 \pm 78 $ & $ 177 \pm 79 $ \\ 
1.225 & 0.11 & $ 188 \pm 76 $ & $ 398 \pm 74 $ & $ 928 \pm 77 $ & $ 1276 \pm 62 $ & $ 120 \pm 64 $ & $ 42 \pm 65 $ & $ 1216 \pm 62 $ & $ 888 \pm 71 $ & $ 539 \pm 75 $ & $ 280 \pm 76 $ \\ 
1.275 & 0.11 & $ 166 \pm 76 $ & $ 379 \pm 74 $ & $ 869 \pm 77 $ & $ 1323 \pm 62 $ & $ 139 \pm 64 $ & $ -8 \pm 65 $ & $ 1282 \pm 62 $ & $ 814 \pm 71 $ & $ 515 \pm 75 $ & $ 247 \pm 76 $ \\ 
1.325 & 0.11 & $ 266 \pm 75 $ & $ 432 \pm 73 $ & $ 904 \pm 76 $ & $ 1357 \pm 62 $ & $ 4 \pm 63 $ & $ 144 \pm 64 $ & $ 1267 \pm 61 $ & $ 925 \pm 70 $ & $ 552 \pm 74 $ & $ 333 \pm 75 $ \\ 
1.375 & 0.12 & $ 189 \pm 81 $ & $ 514 \pm 78 $ & $ 928 \pm 82 $ & $ 1376 \pm 66 $ & $ 131 \pm 67 $ & $ 52 \pm 68 $ & $ 1411 \pm 65 $ & $ 954 \pm 75 $ & $ 461 \pm 79 $ & $ 292 \pm 81 $ \\ 
1.425 & 0.13 & $ 238 \pm 79 $ & $ 532 \pm 76 $ & $ 1198 \pm 79 $ & $ 1431 \pm 64 $ & $ 40 \pm 66 $ & $ 185 \pm 66 $ & $ 1511 \pm 64 $ & $ 1063 \pm 73 $ & $ 605 \pm 77 $ & $ 338 \pm 79 $ \\ 
1.475 & 0.14 & $ 191 \pm 81 $ & $ 527 \pm 79 $ & $ 1068 \pm 82 $ & $ 1460 \pm 66 $ & $ 90 \pm 68 $ & $ 105 \pm 69 $ & $ 1392 \pm 66 $ & $ 1090 \pm 75 $ & $ 580 \pm 80 $ & $ 268 \pm 81 $ \\ 
1.525 & 0.14 & $ 143 \pm 84 $ & $ 478 \pm 81 $ & $ 1048 \pm 85 $ & $ 1429 \pm 69 $ & $ 103 \pm 70 $ & $ 88 \pm 71 $ & $ 1367 \pm 68 $ & $ 943 \pm 77 $ & $ 607 \pm 82 $ & $ 291 \pm 84 $ \\ 
1.575 & 0.15 & $ 367 \pm 88 $ & $ 761 \pm 85 $ & $ 1088 \pm 89 $ & $ 1581 \pm 72 $ & $ 96 \pm 73 $ & $ 125 \pm 74 $ & $ 1503 \pm 71 $ & $ 1107 \pm 81 $ & $ 754 \pm 86 $ & $ 422 \pm 88 $ \\ 
1.625 & 0.16 & $ 359 \pm 93 $ & $ 565 \pm 90 $ & $ 1169 \pm 94 $ & $ 1590 \pm 76 $ & $ 45 \pm 78 $ & $ 181 \pm 79 $ & $ 1593 \pm 75 $ & $ 1142 \pm 86 $ & $ 542 \pm 91 $ & $ 351 \pm 93 $ \\ 
3.6 & 0.17 & $ 1148 \pm 333 $ & $ 3633 \pm 330 $ & $ 4487 \pm 320 $ & $ 4107 \pm 314 $ & $ 661 \pm 324 $ & $ 373 \pm 324 $ & $ 3495 \pm 315 $ & $ 3720 \pm 330 $ & $ 2102 \pm 331 $ & $ 1431 \pm 334 $ \\ 
4.5 & 0.16 & $ 1586 \pm 220 $ & $ 3468 \pm 214 $ & $ 4188 \pm 190 $ & $ 5334 \pm 179 $ & $ 466 \pm 200 $ & $ 757 \pm 200 $ & $ 4904 \pm 183 $ & $ 3626 \pm 214 $ & $ 2575 \pm 217 $ & $ 1057 \pm 221 $ \\ 
\enddata
\vspace{-0.8cm}
\tablecomments{comments}
\end{deluxetable*}

\begin{deluxetable}{lllllll}
\tablecolumns{7}
\tablewidth{0pt}:
\tablecaption{Model Comparison \label{table:models}}
\tablehead{
\colhead{Data} & \colhead{Model} & \colhead{$T_\mathrm{min}$} & \colhead{$T_\mathrm{max}$} & \colhead{$\chi^2$} & \colhead{$\Delta_\mathrm{AIC}$} & \colhead{$\Delta_\mathrm{BIC}$}}
\startdata
WFC3 & Sph. Harmonics & 1214 & 3251 & 1.5 & 0.0 & 0.0 \\
\, & Kinematic & 1980 & 3958 & 1.6 & 11.8 & 11.8 \\
\, & Two Temp. & 0 & 2887 & 1.6 & 47.7 & 43.2 \\
Ch 1 & Sph. Harmonics & 1280 & 3444 & 0.9 & 0.0 & 0.0 \\
\, & Kinematic & 1955 & 3675 & 0.9 & 34.3 & 34.3 \\
\, & Two Temp. & 1429 & 3041 & 0.9 & 10.7 & 17.6 \\
Ch 2 & Sph. Harmonics & 902 & 3786 & 1.1 & 0.0 & 3.1 \\
\, & Kinematic & 1703 & 4276 & 1.1 & 14.35 & 17.41 \\
\, & Two Temp. & 1361 & 3311 & 1.1 & 23.9 & 0.0 \\
\enddata
\vspace{-0.8cm}
\tablecomments{comments}
\end{deluxetable}
